---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 96
)
```

# signs <a href='https://benjaminwolfe.github.io/signs/'><img src='man/figures/logo.svg' align="right" height="139" /></a>

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/BenjaminWolfe/signs.svg?branch=master)](https://travis-ci.org/BenjaminWolfe/signs)
[![Codecov test coverage](https://codecov.io/gh/BenjaminWolfe/signs/branch/master/graph/badge.svg)](https://codecov.io/gh/BenjaminWolfe/signs?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/signs)](https://CRAN.R-project.org/package=signs)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
<!-- badges: end -->

`signs` makes it easy to use typographically accurate minus signs
in plots, markdown, dashboards, or other forms of display.

Ask any typography nut, and they can walk you through the differences
among 4 symbols that look nearly identical:

* the hyphen-minus (-, ASCII 45, next to the zero key)
* the en-dash (`r "\u2013"`, Unicode 2013, Windows Alt Code 0151)
* the em-dash (`r "\u2014"`, Unicode 2014, Windows Alt Code 0150)
* the true minus (`r "\u2212"`, Unicode 2212)

The true minus, when a typeface has one, is _designed for mathematics_.
It has the same length and line thickness as the crossbar of the plus,
and it sits at the same height above the baseline.
By comparison, the hyphen-minus is quite a bit shorter and sits lower.
Hyphens were designed, not for mathematics, but for joining words.
Surprisingly enough, there is no Windows Alt Code for true minuses.

For me, using a Unicode minus is like brewing high-end tea
or wearing my favorite socks.
Maybe nobody else notices, but it still looks good to me.
If you like me enjoy the little things,
this package may grow on you!

## Installation

You can install `signs` from [CRAN](https://cran.r-project.org/package=signs),

``` r
install.packages("signs")
```

or grab the dev version from [GitHub](https://github.com/BenjaminWolfe/signs):

``` r
# install.packages("devtools")
devtools::install_github("BenjaminWolfe/signs")
```

## Example

Using `signs` is simple, especially if you're familiar with functions like
`scales::number()`, `scales::number_format()`,
`scales::comma()`, `scales::comma_format()`,
`scales::percent()`, and `scales::percent_format()`.
It simply provides two new functions to complement these:
`signs::signs()` and `signs::signs_format()`.

See `vignette("signs")` for a full tour of the package.

### Fixed-Width Fonts (indistinguishable)

The difference between minuses is impossible to see with most coding typefaces.
That's not a problem with the package; it's just how fixed-width fonts work.
(How ugly would your code look if characters had different widths?)

```{r text}
library(scales)
library(signs)

x <- seq(-4, 4)

# hyphen-minuses
number(x, accuracy = 1)

# true minuses
signs(x, accuracy = 1)
```

### Plots (distinguishable)

We can see the difference in a plot.

* Points in group 1 are labeled
  with a true Unicode minus glyph: `signs()`.
  So is the y-axis: `signs_format()`.
* Points in group 2 are labeled
  with the traditional ASCII hyphen-minus: `scales::number()`.
* Basic usage is identical.

```{r plots, message = FALSE}
library(dplyr)
library(ggplot2)
library(ggrepel)

theme_set(theme_gray())
theme_update(panel.grid.minor = element_blank())

p <- 
  ggplot(sleep) +
  aes(group, extra) +
  geom_point() +
  xlab("Drug") +
  ylab("Extra Sleep (hours)")

label_hours <- function(mapping) {
  geom_text_repel(
    mapping,
    nudge_x       = -.1,
    direction     = "y",
    segment.size  = .4,
    segment.color = "grey75",
    hjust         = "right"
  )
}

p +
  label_hours(
    mapping = aes(
      label = case_when(
        group == 1 ~ signs(extra, accuracy = .1), # Unicode minuses
        group == 2 ~ number(extra, accuracy = .1) # ASCII minuses
      )
    )
  ) +
  scale_y_continuous(
    limits = c(-4, 6),
    breaks = seq(-4, 6),
    labels = signs_format(accuracy = .1) # Unicode, analogous to number_format() 
  )
```
